Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-17 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Last Revision: small changes to improve readibility

More recently, \citet{howison_2000} and \citet{praud_2005} provided a comprehensive description of the mathematical formulation of immiscible two-phase flows in Hele-Shaw cells. For a Hele-Shaw cell of a given size, flow development depends only on the capillary number, therefore if \textit{$N_{c}$} is too high, \citet{saffman_1959b} \citep[see also][]{saffman_1959,homsy_1987,tabeling_1987} demonstrated that the flow develops to a single steady-state finger which moves throughout the cell with constant velocity. The impact of viscosity ratio and mass flow-rate on fingering development was experimentally (coreflood and micromodels) studied by \citet{Doorwar_2016}. Through numerical (based on pore-network model) experiments, they demonstrated that fingers' dynamics (\ie growth, splitting and coalescence) are also strongly dependent on the domain's aspect-ratio and capillary forces \citep[see][]{Doorwar_2014}, and proposed a new instability dimensionless number to assess breakthrough recovery of a number of experimental data sets.

@@

\medskip 
In this work, a novel CVFEM formulation, previously introduced by \citet{gomes_2017} \citep[see also][]{jackson_2015,salinas2015}, is used to simulate viscous flow instabilities in porous media. The continuity equation is embedded into the pressure equation to enforce mass conservation whilst ensuring that the force balance is preserved. Thus, the main aim of this work is to numerically investigate formation and growth of viscous fingers in heterogeneous porous media through a novel computational multi-fluid dynamics model (embedded in the next-generation flow simulator Fluidity/IC-FERST software). The two main aspects of this work are: (a) qualitative and quantitative validation of the model (Section~\ref{section:results_homo_hete}) for viscous flow instabilities and, (b) impact of mesh resolution on capturing the onset formation and growth of viscous fingers (Section~\ref{section:results_hete_fix_adapt}).

@@

\medskip
A brief description of the numerical formulation and viscous flow instabilities are introduced in Sections~\ref{equations_scheme} and~\ref{section:ViscousInstabilities}, respectively. Model set up and results including initial model-benchmark are presented in Section~\ref{section:results}. Impact of viscosity ratio on the fingers formation is also included in this section. Finally, concluding remarks are presented in Section~\ref{Section:Conclusion}.

>>>>

More recently, \citet{howison_2000} and \citet{praud_2005} provided a comprehensive description of the mathematical formulation of immiscible two-phase flows in Hele-Shaw cells. For a Hele-Shaw cell of a given size, flow development depends only on the capillary number, therefore if \textit{$N_{c}$} is too high, \citet{saffman_1959b} \citep[see also][]{saffman_1959,homsy_1987,tabeling_1987} demonstrated that the flow develops to a single steady-state finger which moves throughout the cell with constant velocity.

@@

\medskip 
The impact of viscosity ratio and flow rate on fingering development was experimentally investigated (with coreflood and micromodels) by \citet{Doorwar_2016}. Through numerical experiments (based on pore-network model), they demonstrated that fingers' dynamics (\ie growth, splitting and coalescence) are also strongly dependent on the domain's aspect-ratio and capillary forces \citep[see][]{Doorwar_2014}, and proposed a new instability dimensionless number to assess breakthrough recovery of an extensive set of experimental data.

Therefore, the main aim of this work is to numerically investigate the onset of viscous instabilities in heterogeneous porous media flows through a novel computational multi-fluid dynamics model~\citep{gomes_2017}. The two main aspects of this work are: (a) qualitative and quantitative validation of the model (Section~\ref{section:results_homo_hete}) for viscous flow instabilities and, (b) impact of mesh resolution on capturing the onset formation and growth of viscous fingers (Section~\ref{section:results_hete_fix_adapt}).

@@

\medskip
A brief description of the numerical formulation and viscous flow instabilities are introduced in Sections~\ref{equations_scheme} and~\ref{section:ViscousInstabilities}, respectively. Model set up and results including initial model-benchmark are presented in Section~\ref{section:results}. Finally, concluding remarks are presented in Section~\ref{Section:Conclusion}.
